---
title: "Getting Started with weathercan2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with weathercan2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

`weathercan2` is an R package for downloading historical weather data from Environment and Climate Change Canada (ECCC) using the MSC GeoMet API. This package provides functions to search for weather stations and download daily, hourly, and monthly weather observations.

This package is inspired by the original `weathercan` package but uses the modern MSC GeoMet API for faster and more reliable data access.

# Installation

```{r installation, eval=FALSE}
# Install from CRAN (when available)
install.packages("weathercan2")

# Install from GitHub
devtools::install_github("yourusername/weathercan2")
```

```{r load-package, eval=FALSE}
library(weathercan2)
```

# Finding Stations

The first step in downloading weather data is finding the station(s) you're interested in. The `stations()` function returns all available climate stations:

```{r all-stations, eval=FALSE}
# Get all stations (this may take a while)
all_stations <- stations()

# View the first few stations
head(all_stations)
```

The resulting tibble contains information about each station:

- `station_name`: The name of the station
- `station_id`: The internal station ID
- `climate_id`: The climate identifier used in API queries
- `prov`: Province/territory code
- `lat`, `lon`: Coordinates
- `elev`: Elevation in meters
- `tz`: Timezone
- `has_daily`, `has_hourly`, `has_monthly`: Boolean indicating data availability
- `has_normals`: Whether climate normals are available

## Searching for Stations

You can search for stations using the `stations_search()` function:

### Search by Name

```{r search-name, eval=FALSE}
# Search for stations with "Kamloops" in the name
kamloops <- stations_search(name = "Kamloops")
print(kamloops)
```

### Search by Province

```{r search-prov, eval=FALSE}
# Search for stations in British Columbia
bc_stations <- stations_search(prov = "BC")
print(bc_stations)
```

### Search by Data Interval

```{r search-interval, eval=FALSE}
# Search for hourly stations
hourly_stations <- stations_search(prov = "BC", interval = "hour")

# Search for daily stations
daily_stations <- stations_search(prov = "ON", interval = "day")
```

### Search by Proximity

```{search-proximity, eval=FALSE}
# Find stations within 20km of a location
nearby <- stations_search(coords = c(50.7, -120.3), dist = 20)
print(nearby)
```

### Search by Climate ID

```{r search-climate-id, eval=FALSE}
# Search for a specific station by climate ID
station <- stations_search(climate_id = "101AE00")
print(station)
```

# Downloading Weather Data

Once you've identified the station(s) you want data from, use `weather_dl()` to download weather observations:

## Download Daily Data

```{r download-daily, eval=FALSE}
# Download daily data for a station
daily_data <- weather_dl(
  climate_id = "101AE00",
  start = "2020-01-01",
  end = "2020-12-31",
  interval = "day"
)

head(daily_data)
```

## Download Hourly Data

```{r download-hourly, eval=FALSE}
# Download hourly data
hourly_data <- weather_dl(
  climate_id = "101AE00",
  start = "2020-06-01",
  end = "2020-06-07",
  interval = "hour"
)

head(hourly_data)
```

## Download for Multiple Stations

```{r download-multiple, eval=FALSE}
# Download data for multiple stations
multi_data <- weather_dl(
  station_ids = c(114, 2, 3),
  start = "2020-01-01",
  end = "2020-01-31",
  interval = "day"
)
```

## Using Station Names

```{r download-name, eval=FALSE}
# Search by station name and download
data <- weather_dl(
  station_names = "Victoria Gonzales",
  start = "2019-01-01",
  end = "2019-12-31",
  interval = "day"
)
```

# Data Structure

The downloaded weather data includes:

- Station metadata (name, ID, location, elevation, timezone)
- Date/time information
- Weather observations (temperature, precipitation, wind, humidity, etc.)
- Quality flags (where available)

For example, daily observations typically include:

- `LOCAL_DATE`: Date of observation
- `MEAN_TEMPERATURE`: Mean daily temperature
- `MIN_TEMPERATURE`, `MAX_TEMPERATURE`: Daily min/max temperatures
- `TOTAL_PRECIPITATION`: Total precipitation
- `TOTAL_RAIN`, `TOTAL_SNOW`: Rain and snow amounts
- `SNOW_ON_GROUND`: Snow depth
- `HEATING_DEGREE_DAYS`, `COOLING_DEGREE_DAYS`: Degree days
- Humidity measurements
- Wind measurements

# Climate Normals

Climate normals are 30-year averages of climate variables. Download them with `normals_dl()`:

```{r normals, eval=FALSE}
# Download current normals (1991-2020)
current_normals <- normals_dl("101AE00")
print(current_normals)

# Download historical normals
normals_1980 <- normals_dl("101AE00", normals_years = "1981-2010")
print(normals_1980)
```

# Complete Workflow Example

Here's a complete example workflow:

```{r workflow, eval=FALSE}
library(weathercan2)
library(dplyr)
library(ggplot2)

# 1. Find stations in a region
stations <- stations_search(prov = "BC", interval = "day")
stations
#> # A tibble: ... x ...
#>   station_name station_id climate_id prov   ...
#>   <chr>            <dbl> <chr>       <chr>  ...
#> 1 VICTORIA GONZALEZ CS   114  1018611    BC     ...

# 2. Select a station
station <- stations[1, ]

# 3. Download data
data <- weather_dl(
  station_ids = station$station_id,
  start = "2020-01-01",
  end = "2020-12-31",
  interval = "day",
  verbose = TRUE
)

# 4. Analyze the data
# View temperature data
data %>%
  select(LOCAL_DATE, MEAN_TEMPERATURE, MAX_TEMPERATURE, MIN_TEMPERATURE) %>%
  head()

# Plot temperature over time
ggplot(data, aes(x = LOCAL_DATE, y = MEAN_TEMPERATURE)) +
  geom_line() +
  labs(
    title = paste("Temperature at", station$station_name),
    x = "Date",
    y = "Mean Temperature (°C)"
  ) +
  theme_minimal()
```

# API Limitations

Be aware of the following limitations when using the MSC GeoMet API:

- Large date ranges may take time to download
- There is a limit of 10,000 records per query (handled automatically by pagination)
- Some historical data may not be available
- API is free and anonymous, but be respectful with large downloads

# Data License

Weather data from Environment and Climate Change Canada is licensed under the [Open Government License - Canada](http://open.canada.ca/en/open-government-licence-canada).

# Getting Help

If you encounter issues or have questions:

1. Check the function documentation with `?function_name`
2. Report issues on GitHub at https://github.com/yourusername/weathercan2/issues
3. Contact ECCC at donneesclimatiquesenligne-climatedataonline@ec.gc.ca
